Manuscript

### 2018-11-21 manuscript.Rmd
update rmd

```{r setup, include=FALSE}
set.seed(1234)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached)
```
cd ~
touch .Renviron
open .Renviron

# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

@@

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name, grouping_var) {
  if(grouping_var == "lineage") {
    grouper <- c("Hugo_Symbol", "group_general_lineage_name")
  }
  else if(grouping_var == "gene") {
    grouper <- "Hugo_Symbol"
  }
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = grouper, data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}

# WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
#   druglist <- as.character(unique(melt_file$Drug))
#   sig_df <- NULL
#   for(i in 1:length(druglist)) {
#     print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
#     flush.console()
#     melted <- filter(melt_file, Drug == druglist[i])
#     use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
#     use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
#     use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
#     use_data$AUC <- as.numeric(use_data$AUC)
#     use_data <- filter(use_data, !is.na(use_data$AUC))
#     use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
#     use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
#     use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
#     use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
#     
#     sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
#     sig <- adj_signif(sig)
#     sig <- sig[order(sig$p),]
#     sig <- filter(sig, p <= 0.05)
#     sig$Drug <- druglist[i]
#     sig$Dataset <- toupper(data_name)
#     sig_df <- rbind(sig_df, sig)
#   }
#   return(sig_df)
# }
```

@@

  y_value <- ifelse(ticker == 3, 10, max(use_data$n) + 1)

  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                          Test_Results_p = c(ccle_lin_p_ttest_res, ctrp_lin_p_ttest_res, gdsc_lin_p_ttest_res),
                          t_statistic = c(round(ccle_lin_p_ttest$statistic, 3), round(ctrp_lin_p_ttest$statistic, 3), round(gdsc_lin_p_ttest$statistic, 3)),
                          df = c(round(ccle_lin_p_ttest$parameter, 3), round(ctrp_lin_p_ttest$parameter, 3), round(gdsc_lin_p_ttest$parameter, 3)),
                          p_value = c(round(ccle_lin_p_ttest$p.value, 3), round(ctrp_lin_p_ttest$p.value, 3), round(gdsc_lin_p_ttest$p.value, 3)),
                          group_general_lineage_name = lineage_name)

  plot_point <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, y_value)) +
    scale_x_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    scale_y_continuous(breaks = pretty_breaks()) +
    geom_point(mapping = aes(label = Drug_Gene, color = InG2P, x = p, y = n), alpha = 0.5, position = position_jitter(width = 0, height = 0.15)) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    geom_text_repel(data = subset(use_data, InG2P == "Yes"), mapping = aes(x = p, y = n, label = Drug_Gene), size = 3, segment.size = 0.2, segment.color = "grey50") +
    geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value, label = Test_Results_p), nudge_y = 0.1) +
    labs(x = "Drug-gene association p-value between mutation groups", y = "Number of level A G2P associations") +
    theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())

  plot_box <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(ylim = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    geom_violin(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.5) +
    geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.75, width = 0.25) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    coord_flip() +
    labs(x = "In G2P?", title = toupper(lineage_name)) +
    theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

  plot <- subplot(plot_box, plot_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>% layout(title = toupper(lineage_name), margin = list(t = 75, l = 75), xaxis5 = list(title = "Gene-drug association p-value between mutation groups"), yaxis = list(title = "Level A G2P\nassociation", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

  saveWidgetFix(widget = plot, file = paste0("./plots/manuscript/plotlys_mut_by_lineage/", lineage_name, "_mut_plotly.html"))

@@

```{r}
ccl_info <- read.delim("./data_munging/18Q4/DepMap-2018q4-celllines_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)

# From figshare
crispr_meta <- read.delim("./data_munging/18Q4/sample_info_18Q4_crispr_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

@@

Merged mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wiltype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

@@

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./data_munging/18Q4/public_18Q4_gene_cn_20181115.csv.gz", sep = ",", check.names = FALSE, header = TRUE)

# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(cn) <- gsub(" .*", "", colnames(cn))

# Melt
cn_melt <- melt(data = cn, id.vars = "DepMap_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")

saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q4.rds", compress = "xz")
```

@@

18Q3, Gene expression data (Reads Per Kilobase of transcript, per Million mapped reads, RPKM).
```{r, eval=FALSE}
ge_18Q3 <- read.delim("./../htscreen_lineages_giant_files/CCLE_DepMap_18q3_RNAseq_RPKM_20180718.gct.gz", skip = 2, header = TRUE, sep = "\t", check.names = FALSE)

# Edit columns
ge_18Q3$Name <- NULL
colnames(ge_18Q3)[1] <- "Hugo_Symbol"

# Melt
ge_melt_18Q3 <- melt(data = ge_18Q3, id.vars = "Hugo_Symbol", measure.vars = colnames(ge_18Q3[2:ncol(ge_18Q3)]), value.name = "RPKM")
## Split variable column
ge_melt_18Q3 <- with(ge_melt_18Q3, cbind(Hugo_Symbol, colsplit(variable, pattern = " ", names = c("CCLE_Name", "DepMap_ID")), RPKM))
## Remove parentheses around Broad IDs
ge_melt_18Q3$DepMap_ID <- gsub("\\(|\\)", "", ge_melt_18Q3$DepMap_ID)

saveRDS(ge_melt_18Q3, "./../htscreen_lineages_giant_files/ge_melt_18Q3.rds", compress = "xz")
```

18Q4, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge_18Q4 <- read.delim("./../htscreen_lineages_giant_files/CCLE_depMap_18Q4_TPM_v2_20181115.csv.gz", header = TRUE, sep = ",", check.names = FALSE)

# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(ge_18Q4) <- gsub(" .*", "", colnames(ge_18Q4))
colnames(ge_18Q4)[1] <- "DepMap_ID"

# Melt
ge_melt_18Q4_all <- melt(ge_18Q4, id.vars = "DepMap_ID", measure.vars = colnames(ge_18Q4)[2:ncol(ge_18Q4)], variable.name = "Hugo_Symbol", value.name = "TPM")
ge_melt_18Q4 <- unique(ge_melt_18Q4)

saveRDS(ge_melt_18Q4, "./data_munging/rds/ge_melt_18Q4.rds", compress = "xz")
```

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_withGroupedLineages.csv", sep = "\t", header = TRUE, row.names = 1)
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)
g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_drug_genes <- unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)])
g2p_genes <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[2]))
g2p_drugs <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[1]))

g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

@@

```{r}
crispr_data <- readRDS("./../htscreen_lineages_giant_files/crispr_data_18Q4.rds")

crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
```

@@

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
ge_filt_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% g2p_genes)
# ge_filt_18Q4 <- filter(ge_melt_18Q4, Hugo_Symbol %in% g2p_genes)
cn_filt <- filter(cn_melt, Hugo_Symbol %in% g2p_genes)

>>>>

```{r setup, include=FALSE}
set.seed(1234)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached)
```
cd ~
touch .Renviron
open .Renviron
# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

@@

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name, grouping_var) {
  if(grouping_var == "lineage") {
    grouper <- c("Hugo_Symbol", "group_general_lineage_name")
  }
  else if(grouping_var == "gene") {
    grouper <- "Hugo_Symbol"
  }
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = grouper, data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}
# WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
#   druglist <- as.character(unique(melt_file$Drug))
#   sig_df <- NULL
#   for(i in 1:length(druglist)) {
#     print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
#     flush.console()
#     melted <- filter(melt_file, Drug == druglist[i])
#     use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
#     use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
#     use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
#     use_data$AUC <- as.numeric(use_data$AUC)
#     use_data <- filter(use_data, !is.na(use_data$AUC))
#     use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
#     use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
#     use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
#     use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
#     
#     sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
#     sig <- adj_signif(sig)
#     sig <- sig[order(sig$p),]
#     sig <- filter(sig, p <= 0.05)
#     sig$Drug <- druglist[i]
#     sig$Dataset <- toupper(data_name)
#     sig_df <- rbind(sig_df, sig)
#   }
#   return(sig_df)
# }
```

@@

  y_value <- ifelse(ticker == 3, 10, max(use_data$n) + 1)
  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                          Test_Results_p = c(ccle_lin_p_ttest_res, ctrp_lin_p_ttest_res, gdsc_lin_p_ttest_res),
                          t_statistic = c(round(ccle_lin_p_ttest$statistic, 3), round(ctrp_lin_p_ttest$statistic, 3), round(gdsc_lin_p_ttest$statistic, 3)),
                          df = c(round(ccle_lin_p_ttest$parameter, 3), round(ctrp_lin_p_ttest$parameter, 3), round(gdsc_lin_p_ttest$parameter, 3)),
                          p_value = c(round(ccle_lin_p_ttest$p.value, 3), round(ctrp_lin_p_ttest$p.value, 3), round(gdsc_lin_p_ttest$p.value, 3)),
                          group_general_lineage_name = lineage_name)
  plot_point <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, y_value)) +
    scale_x_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    scale_y_continuous(breaks = pretty_breaks()) +
    geom_point(mapping = aes(label = Drug_Gene, color = InG2P, x = p, y = n), alpha = 0.5, position = position_jitter(width = 0, height = 0.15)) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    geom_text_repel(data = subset(use_data, InG2P == "Yes"), mapping = aes(x = p, y = n, label = Drug_Gene), size = 3, segment.size = 0.2, segment.color = "grey50") +
    geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value, label = Test_Results_p), nudge_y = 0.1) +
    labs(x = "Drug-gene association p-value between mutation groups", y = "Number of level A G2P associations") +
    theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
  plot_box <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(ylim = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    geom_violin(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.5) +
    geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.75, width = 0.25) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    coord_flip() +
    labs(x = "In G2P?", title = toupper(lineage_name)) +
    theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))
  plot <- subplot(plot_box, plot_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>% layout(title = toupper(lineage_name), margin = list(t = 75, l = 75), xaxis5 = list(title = "Gene-drug association p-value between mutation groups"), yaxis = list(title = "Level A G2P\nassociation", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))
  saveWidgetFix(widget = plot, file = paste0("./plots/manuscript/plotlys_mut_by_lineage/", lineage_name, "_mut_plotly.html"))

@@

```{r}
ccl_info <- read.delim("./data_munging/18Q4/DepMap-2018q4-celllines_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)
# From figshare
crispr_meta <- read.delim("./data_munging/18Q4/sample_info_18Q4_crispr_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

@@

Merged mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wiltype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

@@

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./data_munging/18Q4/public_18Q4_gene_cn_20181115.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(cn) <- gsub(" .*", "", colnames(cn))
# Melt
cn_melt <- melt(data = cn, id.vars = "DepMap_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")
saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q4.rds", compress = "xz")
```

@@

18Q3, Gene expression data (Reads Per Kilobase of transcript, per Million mapped reads, RPKM).
```{r, eval=FALSE}
ge_18Q3 <- read.delim("./../htscreen_lineages_giant_files/CCLE_DepMap_18q3_RNAseq_RPKM_20180718.gct.gz", skip = 2, header = TRUE, sep = "\t", check.names = FALSE)
# Edit columns
ge_18Q3$Name <- NULL
colnames(ge_18Q3)[1] <- "Hugo_Symbol"
# Melt
ge_melt_18Q3 <- melt(data = ge_18Q3, id.vars = "Hugo_Symbol", measure.vars = colnames(ge_18Q3[2:ncol(ge_18Q3)]), value.name = "RPKM")
## Split variable column
ge_melt_18Q3 <- with(ge_melt_18Q3, cbind(Hugo_Symbol, colsplit(variable, pattern = " ", names = c("CCLE_Name", "DepMap_ID")), RPKM))
## Remove parentheses around Broad IDs
ge_melt_18Q3$DepMap_ID <- gsub("\\(|\\)", "", ge_melt_18Q3$DepMap_ID)
saveRDS(ge_melt_18Q3, "./../htscreen_lineages_giant_files/ge_melt_18Q3.rds", compress = "xz")
```

18Q4, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge_18Q4 <- read.delim("./../htscreen_lineages_giant_files/CCLE_depMap_18Q4_TPM_v2_20181115.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(ge_18Q4) <- gsub(" .*", "", colnames(ge_18Q4))
colnames(ge_18Q4)[1] <- "DepMap_ID"
# Melt
ge_melt_18Q4_all <- melt(ge_18Q4, id.vars = "DepMap_ID", measure.vars = colnames(ge_18Q4)[2:ncol(ge_18Q4)], variable.name = "Hugo_Symbol", value.name = "TPM")
ge_melt_18Q4 <- unique(ge_melt_18Q4)
saveRDS(ge_melt_18Q4, "./data_munging/rds/ge_melt_18Q4.rds", compress = "xz")
```

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_withGroupedLineages.csv", sep = "\t", header = TRUE, row.names = 1)
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)
g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))
g2p_drug_genes <- unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)])
g2p_genes <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[2]))
g2p_drugs <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[1]))
g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

@@

```{r}
crispr_data <- readRDS("./../htscreen_lineages_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)
# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
```

@@

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
ge_filt_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% g2p_genes)
# ge_filt_18Q4 <- filter(ge_melt_18Q4, Hugo_Symbol %in% g2p_genes)
cn_filt <- filter(cn_melt, Hugo_Symbol %in% g2p_genes)
# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "ActArea")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Wildtype", ccle_data$Mutation_Status_Nonsilent)
ccle_data <- merge(ccle_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# ccle_data <- merge(ccle_data, ge_filt_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data <- merge(ccle_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$ActArea <- as.numeric(ccle_data$ActArea)
# CCLE AUC is actually AOC
# assuming scale is 0-8, transform to range between 0 and 1
# get AUC by 1 - AOC
ccle_data$AUC <- 1 - (ccle_data$ActArea / 8)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$AUC_zscore <- (ccle_data$AUC - mean(ccle_data$AUC)) / sd(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Wildtype", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q4_g2pgenes.rds", compress = "xz")
# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Wildtype", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data <- merge(ctrp_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# ctrp_data <- merge(ctrp_data, ge_filt_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data <- merge(ctrp_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
